---
title: "Univaf Sandbox"
date: "Last compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, message=F, warning=F, cache=T, cache.lazy=F)
source("shared.R")
source("shared_map.R")


start_date = '2021-05-19'
end_date   =' 2021-07-15'

dloc = read_location_data()
dav  = read_availability_slot_data() %>%
  filter(slot_time > as.POSIXct(paste0(start_date, ' 00:00:00'), 'UTC'),
         slot_time < as.POSIXct(paste0(end_date  , ' 00:00:00'), 'UTC')) %>%
  mutate(ds=as.Date(slot_time))
```


In this report we take cursory look at slot-level COVID appointment availability data as gathered by USDR's [UNIVAF](http://getmyvax.org/docs/) project. The main question we're looking at is whether this kind of data can be used to identify issues in vaccine access by different demographics. Another question is whether this can help us understand the role of barriers to access, as opposed to more ideological barriers, can help us understand trailing vaccination rates.


* This report pertains to *slot-level* appointment data only. We often don't have slot-level data, so this covers a subset of appointments. Of the `r nrow(dloc)` locations in our national DB, `r length(unique(dav$id))` have slot-level data.
* We take as the interval a slot is available for as the time between the first time we saw it being available to the last time.
* 'work time' is defined as Mon-Fri 9am-5pm.
* For slots we compute the following statistics:
    1. how many slots are ever seen to be available for that hour
    2. how long (in number of hours) are they available for on average
    3. what is the average _latest_ time (in number of hours ahead for the appointment time) you can still book the slot




## Number of slots per state per day

```{r, fig.height=7, fig.width=10}
dav %>% 
   group_by(state, ds) %>% summarize(n=n()) %>%
  ggplot(aes(ds, n)) + geom_line() +
    scale_x_date("Date", limits=as.Date(c(start_date, end_date))) +
    scale_y_continuous("Total number of slots (k)", labels=big_numbers) +
    my_theme() + facet_wrap(~state, scales = 'free_y')
```

## Slot stats

```{r, fig.height=4, fig.width=6}
dg = dav %>%
  group_by(ds, dow) %>%
  summarize(n=n(), p_booked=mean(booked), range=mean(range, na.rm=T), lead=mean(last_time_ahead, na.rm=T))

Rmisc::multiplot(
  ggplot(dg, aes(ds, n)) + geom_line() + geom_point() + geom_point(data=dg %>% filter(dow > 4), color='red') +
    xlab("Appointment Date") + scale_y_continuous("Number of slots", limits=c(0, NA), expand=expand_scale(mult=c(0,.1))) + my_theme(),
  ggplot(dg, aes(ds, range)) + geom_line() + geom_point() + geom_point(data=dg %>% filter(dow > 4), color='red') +
    xlab("Appointment Date") + scale_y_continuous("Avg hours slots available for", limits=c(0, NA), breaks=c(0:6)*24, expand=expand_scale(mult=c(0,.1))) + my_theme(),
  ggplot(dg, aes(ds, p_booked)) + geom_line() + geom_point() + geom_point(data=dg %>% filter(dow > 4), color='red') +
    xlab("Appointment Date") + scale_y_continuous("Share slots \"booked\"", limits=c(0, NA), expand=expand_scale(mult=c(0,.1))) + my_theme(),
  ggplot(dg, aes(ds, lead)) + geom_line() + geom_point() + geom_point(data=dg %>% filter(dow > 4), color='red') +
    xlab("Appointment Date") + scale_y_continuous("Avg hours before slots booked", limits=c(0, NA), expand=expand_scale(mult=c(0,.1))) + my_theme(),
  cols=2)
```


by state
```{r, fig.height=6, fig.width=9}
dgs = dav %>%
  group_by(ds, dow, state) %>%
  summarize(n=n(), p_booked=mean(booked), range=mean(range, na.rm=T), lead=mean(last_time_ahead, na.rm=T))

ggplot(dgs, aes(ds, range)) + geom_line() + geom_point(data=dgs %>% filter(dow > 4), color='red') +
    xlab("Appointment Date") + scale_y_continuous("Avg hours slots available for", breaks=c(0:6)*24, expand=expand_scale(mult=c(0,.1))) + my_theme() +
  coord_cartesian(ylim=c(0, 140)) + facet_wrap(~state)

ggplot(dgs, aes(ds, lead)) + geom_line() + geom_point(data=dgs %>% filter(dow > 4), color='red') +
  xlab("Appointment Date") + scale_y_continuous("Avg hours before slots booked", expand=expand_scale(mult=c(0,.1))) + my_theme() +
  coord_cartesian(ylim=c(0, 40)) +
  facet_wrap(~state)
```

## Slots per person by ToW by State

```{r, fig.height=6, fig.width=9}
dav %>%
  #filter(state %in% c("SD",'ND')) %>%
  filter(slot_time > as.POSIXct('2021-05-21 00:00:00', 'UTC'),
         slot_time < as.POSIXct('2021-07-16 00:00:00', 'UTC')) %>%
  select(-county) %>% inner_join(dzip %>% select(zip, fips=county_fips), by='zip') %>%
  group_by(ds, dow, hod, fips, state) %>% summarize(n_slots=n()) %>%
  left_join(dcou %>% select(fips, population), by='fips') %>%
  group_by(dow, hod, state) %>% summarize(stat=mean(n_slots/population*1000)) %>%
  mutate(stat=ifelse(stat > 0.25, 0.25, stat)) %>%
  dow_plot("Average number of slots/person", wrap_by='state', wrap_cols = 10)
```

## Visualization of range of `checked_at` time

* sample 10 appointments per day (5 from walgreens and 5 other ones)
* Mondays are highlighted

```{r, fig.height=5, fig.width=8}
dx <- dav %>%
  #filter(slot_time > as.POSIXct('2021-05-23 00:00:00', 'UTC'),
  #       slot_time < as.POSIXct('2021-06-21 00:00:00', 'UTC')) %>%
  mutate(wg=provider=='walgreens') %>%
  group_by(ds, wg) %>% sample_n(5) %>% ungroup() %>%
  arrange(slot_time) %>%
  mutate(row_id=row_number())

ggplot(dx, aes(x=slot_time, y=row_id)) + geom_point() + 
    geom_point(data=dx %>% filter(dow==0), color='red') +
    #geom_point(aes(color=factor(dow, labels=c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")))) +
    geom_segment(aes(x=min, xend=max, yend=row_id), color='grey') + 
    #geom_segment(aes(x=min, xend=max, yend=row_id, color=provider)) + 
    scale_x_datetime("Time", limits=c(as.POSIXct('2021-05-20'), NA)) +
    my_theme() + theme(axis.title.y=element_blank(),
                       axis.text.y=element_blank(),
                       axis.ticks.y=element_blank())
```

* Wallgreens is overrepresented and releases their appointments on Friday apparently.
* some appointments are availale after the slot time


## Availability by vulnerability (national, by zip)

We can compare appointment availability by the demographics of the area that the distribution location is in. The most granual level is the zip-code. Of course, people can travel to locations outside of their zip-code, and so a better analysis would, for every zip code, take an average of the nearby locations weighted by their distance.

```{r}
stats = dav %>% 
  filter(slot_time > as.POSIXct('2021-05-19 00:00:00', 'UTC'),
         slot_time < as.POSIXct('2021-06-24 00:00:00', 'UTC')) %>%
    group_by(zip) %>%
  summarize(n_locs=n_distinct(id), n_slots=n(), n_slots_weekend=sum(ifelse(type=='weekend', 1, 0)),
            hours_available=mean(range), last_available=mean(last_time_ahead)) %>%
  inner_join(dzip) %>%
  mutate(slots_per_person=n_slots/total, slots_per_person_weekend=n_slots_weekend/total)
```

We do univariate correlations between vulnerability indices like SVI and availability statistics. In this plot, the points represent individual zip-codes. The red line is a linear fit of the relationship between the vulnerability statistic and the availbility statistic.

```{r, fig.height=5, fig.width=10}
Stats = stats %>%
  select(state, zip, p_black, svi_socioeconomic, svi_household, svi_minority, svi_housing, slots_per_person, hours_available, last_available) %>%
  mutate(slots_per_person=log(slots_per_person), p_black=log(p_black)) %>%
  gather('var_x', 'val_x', -state, -zip, -hours_available, -last_available, -slots_per_person) %>% filter(!is.na(val_x)) %>%
  gather('var_y', 'val_y', -zip, -state, -var_x, -val_x) %>% filter(!is.na(val_y)) %>%
  filter(!(var_y=='hours_available' & val_y > 200)) %>%
  filter(!(var_y=='slots_per_person' & val_y < -8)) %>%
  filter(!(var_y=='last_available' & val_y > 125)) %>%
  mutate(var_x=factor(var_x, levels=c('p_black','svi_socioeconomic','svi_household','svi_minority','svi_housing'),
                     labels=c('Share Black (log)', 'Socioconomic SVI', 'Household SVI', 'Minority SVI', 'Housing SVI')),
         var_y=factor(var_y, levels=c('slots_per_person','hours_available','last_available'),
                      labels=c("Slots/person (log)","Hours Available","Last Available")))

ggplot(Stats, aes(val_x, val_y)) + geom_point(alpha=0.5) +
  #geom_smooth(method='lm', aes(color=S), alpha=0) +
  geom_smooth(data=Stats, method='lm', alpha=0) +
  my_theme() + theme(axis.title.x=element_blank(), axis.title.y=element_blank()) + facet_grid(var_y ~ var_x, scales='free')
```

Note that for 'last available' (min lead time), less is better.
For some indices, we see a **slight negative correlation between vulnerability and availability**.

*Note*: because we don't have vaccination data on the zip-code level, we can't (unlike below for counties) adjust slots/person by coverage, nor can we plot the SVI statistics against vaccination rates on a zip-code level.


We can combine these factors by running linear regressions for the outcome variable we've looked at so far (fixed effects for state not shown):

```{r, eval=T}
stargazer::stargazer(
  lm(slots_per_person ~ log(total) + p_black + svi_socioeconomic + svi_household + svi_minority + svi_housing + state, data=stats),
  lm(hours_available ~ log(total) + p_black + svi_socioeconomic + svi_household + svi_minority + svi_housing + state, data=stats),
  lm(last_available ~ log(total) + p_black + svi_socioeconomic + svi_household + svi_minority + svi_housing + state, data=stats),
  covariate.labels = c("Population (log)", "Share Black", "Socioeonomic SVI", "Household SVI", "Minority SVI", "Housing SVI", "Intercept"),
  type='text', no.space = T, keep.stat = c('n','adj.rsq'),
  omit='state'
)
```

* Unlike in the next section, we can't account for our coverage by dividing slots/person by bookings/vaccinations, so slots/person is at least partially a measure of coverage in this regression and shouldn't be taken very seriously.
* Accounting for all five measures of vulnerability, zip-codes with a higher percentage Black residents have better booking time-based access, while zipcodes that are high on any of the SVI metrics have worse access (shorter booking windows and less time to book).


```{r, eval=F}
stargazer::stargazer(
  lm(slots_per_person ~ svi_socioeconomic*state, data=stats %>% filter(!is.na(slots_per_person), !is.infinite(slots_per_person))),
  lm(hours_available ~ svi_socioeconomic*state, data=stats),
  lm(last_available ~ svi_socioeconomic*state, data=stats),
  #covariate.labels = c("Population (log)", "Share Black", "Socioeonomic SVI", "Household SVI", "Minority SVI", "Housing SVI", "Intercept"),
  type='text', no.space = T, keep.stat = c('n','adj.rsq')
)

CT,IA,FL,IL,MA,MT

ggplot(stats, aes(svi_socioeconomic, hours_available)) + geom_point(shape='.') + geom_smooth(method='lm')


ggplot(Stats %>% filter(state=='CT'), aes(val_x, val_y)) + geom_point() +
  geom_smooth(method='lm', aes(color='CT'), alpha=0) +
  geom_smooth(data=Stats, method='lm', aes(color='US'), alpha=0) +
  my_theme() + theme(axis.title.x=element_blank(), axis.title.y=element_blank()) + facet_grid(var_y ~ var_x, scales='free')
```

## Does availability predict vaccinations? (national, by county)

* data is for 4 weeks: 2021-05-24 to 2021-06-20
* vaccination counts from covid act now

```{r}
dvax = make_vax('2021-05-24', '2021-06-21')

dav  = read_availability_slot_data() %>%
  filter(slot_time > as.POSIXct('2021-05-23 00:00:00', 'UTC'),
         slot_time < as.POSIXct('2021-06-21 00:00:00', 'UTC'))
dagg = make_regression_data(dav, dzip, dvax, dcou, dpol)
```

How does appointment availability relate to vaccinations? Unfortunately we don't have vaccination data on the zip-code level, so we'll do this analysis on a county level. Vaccination counts by county by day are pulled from [CovicActNow.org](https://covidactnow.org/). The outcome we look at is the _number of vaccinations_ that happened in the two weeks that we're looking at. We run a simple linear regression with counties as units:

```{r}
stargazer::stargazer(
  lm(n_vax/population ~ state + slots_per_person_workday + slots_per_person_weekend + slots_per_person_evening
                              + avg_range  + avg_ahead, data=dagg),
  lm(n_vax/population ~ state + hesitant + p_rep_votes + old_vax_rate
                              + slots_per_person_workday + slots_per_person_weekend + slots_per_person_evening
                              + avg_range + avg_ahead, data=dagg),
  lm(log(n_vax) ~ state + log(population), data=dagg),
  lm(log(n_vax) ~ state + log(population) 
                        + slots_per_person_workday + slots_per_person_weekend + slots_per_person_evening
                        + avg_range + avg_ahead, data=dagg),
  lm(log(n_vax) ~ state + log(population) + hesitant + p_rep_votes + old_vax_rate
                        + slots_per_person_workday + slots_per_person_weekend + slots_per_person_evening
                        + avg_range + avg_ahead , data=dagg),
  covariate.labels = c("% hesitant (CDC)", "Trump vote share", "Prior vax rate", 
                       "Slots/Person (workday)", "Slots/Person (weekend)", "Slots/Person (evenings)", "Avg range (hrs)", "Avg lead time (hrs)",
                       "Population (log)", "Intercept"),
  type='text', no.space = T, keep.stat = c('n','adj.rsq'),
  omit='state'
)
```


* **Availability stats are predictive of the number of vaccinations**, and explain about 2% of the variation depending on the specification.
* There are two specifications of the outcome: vaccinations per individual (aka the difference in vaccination rate over the three weeks), and the log number of vaccinations (with the log population as a covariate).
* Hesitancy (taken from [this](https://data.cdc.gov/Vaccinations/Vaccine-Hesitancy-for-COVID-19-County-and-local-es/q9mh-h2tw/data) CDC survey) and the prior vaccination rate are strongly correlated with vaccinations, but don't change the outcomes.




### Week over week

Now, week over week (only two weeks since we don't have vaccination numbers for this week until end of week):

```{r}
dvax2 = read_csv(paste0(data_folder, "/misc/vaccination rates/counties.timeseries.csv")) %>%
  select(date, fips, state, vax=actuals.vaccinationsCompleted, old_vax_rate=metrics.vaccinationsCompletedRatio)

dvaxCW = dvax2 %>%
  mutate(w=isoweek(date)) %>% filter(year(date) == 2021, w > 19, w < 27) %>%
  group_by(fips, w) %>% summarize(max_vax=max(vax, na.rm=T), old_vax_rate=min(old_vax_rate, na.rm=T)) %>%
  group_by(fips) %>% arrange(w) %>% mutate(n_vax=max_vax-lag(max_vax)) %>% ungroup() %>% arrange(fips, w) %>%
  filter(is.finite(n_vax), n_vax > 0) %>%
  mutate(w = w - 1)  # join to the week before!

daggW <- dav %>% select(-county) %>%
  inner_join(dzip %>% select(zip, fips=county_fips), by='zip') %>%
  group_by(state, fips, w) %>%
  summarize(
    n_locations=n_distinct(id),
    n_slots=n(),
    n_slots_workday=sum(ifelse(type=='workday', 1, 0)),
    n_slots_weekend=sum(ifelse(type=='weekend', 1, 0)),
    n_slots_evening=sum(ifelse(type=='evening', 1, 0)),
    avg_range=mean(range),
    avg_range_workday=mean(ifelse(type=='workday', range, NA), na.rm=T),
    avg_range_weekend=mean(ifelse(type=='weekend', range, NA), na.rm=T),
    avg_range_evening=mean(ifelse(type=='evening', range, NA), na.rm=T),
    avg_ahead=mean(last_time_ahead),
    avg_ahead_workday=mean(ifelse(type=='workday', last_time_ahead, NA), na.rm=T),
    avg_ahead_weekend=mean(ifelse(type=='weekend', last_time_ahead, NA), na.rm=T),
    avg_ahead_evening=mean(ifelse(type=='evening', last_time_ahead, NA), na.rm=T),
    n_booked=sum(booked)
  ) %>% ungroup() %>%
  inner_join(dvaxCW, by=c('fips', 'w')) %>%
  inner_join(dcou %>% select(fips, population, vax_rate, hesitant), by='fips') %>%
    mutate(
      slots_per_person=n_slots/population,
      slots_per_person_workday=n_slots_workday/population,
      slots_per_person_weekend=n_slots_weekend/population,
      slots_per_person_evening=n_slots_evening/population,
      book_per_vax=n_booked/n_vax
    ) %>%
  filter(!is.na(fips)) %>%
  # adjust for coveragae
  mutate(
    slots_per_person = slots_per_person/book_per_vax,
    slots_per_person_workday = slots_per_person_workday/book_per_vax,
    slots_per_person_weekend = slots_per_person_weekend/book_per_vax,
    slots_per_person_evening = slots_per_person_evening/book_per_vax
  ) %>%
  filter(is.finite(slots_per_person))
```

```{r}
stargazer::stargazer(
  lm(n_vax/population ~ state, data=daggW),
  lm(n_vax/population ~ slots_per_person + avg_range + avg_ahead + state, data=daggW),
  #lm(n_vax/population ~ slots_per_person_workday + slots_per_person_weekend + slots_per_person_evening
  #                          + avg_range_workday + avg_range_weekend + avg_range_evening
  #                          + avg_ahead_workday + avg_ahead_weekend + avg_ahead_evening
  #                      + state, data=daggW),
  lm(n_vax/population ~ hesitant + old_vax_rate + state, data=daggW),
  lm(n_vax/population ~ hesitant + old_vax_rate + slots_per_person + avg_range + avg_ahead + state, data=daggW),
  lm(log(n_vax) ~ state + log(population), data=daggW),
  lm(log(n_vax) ~ slots_per_person + avg_range + avg_ahead + state + log(population), data=daggW),
  lm(log(n_vax) ~ hesitant + old_vax_rate + state + log(population), data=daggW),
  lm(log(n_vax) ~ hesitant + old_vax_rate + slots_per_person + avg_range + avg_ahead + state + log(population), data=daggW),
  covariate.labels = c("Slots/Person", "Average range (hrs)", "Average time ahead (hrs)", "% hesitant (CDC survey)", "Prior vax rate", "Population (log)", "Intercept"),
  omit='state',
  type='text', no.space = T, keep.stat = c('n','adj.rsq')
)
```

* avg range negative now
* explained variance is 5-12% now.


## Maps

```{r map_data_state}
dmap <- dav %>%
  group_by(state) %>%
  summarize(n_slots=n(), range=mean(range, na.rm=T), n_booked=sum(booked), n_weekend=sum(ifelse(type=='weekend', 1, 0))) %>% ungroup() %>%
  left_join(dvax %>% group_by(state) %>% summarize(n_vax=sum(n_vax, na.rm=T)), by=c('state')) %>%
  left_join(dcou %>% group_by(state) %>% summarize(population=sum(population)), by=c('state')) %>%
  mutate(slots_per_person_raw=n_slots/population, book_per_vax=n_booked/n_vax, weekend=n_weekend/n_slots, weekend2=(n_weekend/2) / (n_slots/7)) %>%
  mutate(slots_per_person = slots_per_person_raw/book_per_vax) %>% #filter(book_per_vax < 1, book_per_vax > 0.1) %>%
  #filter(is.finite(slots_per_person)) %>%
  inner_join(dcou %>% mutate(fips=str_sub(fips, 1, 2)) %>% select(fips, state) %>% distinct())
```

```{r maps_state, fig.height=5, fig.width=8}
plot_map(dmap %>% select(fips, stat=slots_per_person) %>% mutate(stat=ifelse(stat>0.10, 0.10, stat)), title='Slots/Person', layer='state')
plot_map(dmap %>% select(fips, stat=range)  %>% mutate(stat=ifelse(stat<40, 40, ifelse(stat>90, 90, stat))), title='Avg Range', layer='state')
plot_map(dmap %>% select(fips, stat=weekend) %>% mutate(stat=ifelse(stat<0.12, 0.12, stat)), title='Weekend slots/slot', layer='state')
plot_map(dmap %>% select(fips, stat=weekend2) %>% mutate(stat=ifelse(stat<0.4, 0.4, stat)), title='Relative prominence of weekend slots', layer='state')
```


```{r map_data_county}
# our stats
dmap <- dav %>%
  select(-county) %>% inner_join(dzip %>% select(zip, fips=county_fips), by='zip') %>%
  group_by(state, fips) %>%
  summarize(n_slots=n(), range=mean(range, na.rm=T), n_booked=sum(booked), n_weekend=sum(ifelse(type=='weekend', 1, 0))) %>% ungroup() %>%
  left_join(dvax, by=c('fips', 'state')) %>%
  left_join(dcou %>% select(fips, population, vax_rate), by='fips') %>%
  mutate(slots_per_person_raw=n_slots/population, book_per_vax=n_booked/n_vax, weekend=n_weekend/n_slots) %>%
  mutate(slots_per_person = slots_per_person_raw/book_per_vax) %>%
  filter(!is.na(fips), is.finite(slots_per_person))
```

```{r maps_county, fig.height=5, fig.width=8}
plot_map(dmap %>% select(fips, stat=slots_per_person) %>% mutate(stat=ifelse(stat>0.3, 0.3, stat)), title='Slots/Person')
plot_map(dmap %>% select(fips, stat=range) %>% mutate(stat=ifelse(stat>100, 100, stat)), title='Avg Range')
plot_map(dmap %>% select(fips, stat=weekend) %>% mutate(stat=ifelse(stat>0.3, 0.3, ifelse(stat<0.1, 0.1, stat))), title='Weekend Slots/Slots')
```

